### 2019-09-14 paper/poster/poster.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% baposter Landscape Poster
% LaTeX Template
% Version 1.0 (15/5/13)
%
% Created by:
% Brian Amberg (baposter@brian-amberg.de)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[a0paper,portrait,fontscale=0.35, margin=2cm]{baposter}

\usepackage[utf8]{inputenc}
\usepackage[font=normalsize,labelfont=bf]{caption} % Required for specifying captions to tables and figures
\captionsetup{justification=centering}
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{lmodern}
\usepackage{relsize} % Used for making text smaller in some places

% Other packages
\usepackage{tabularx}
\def\tabularxcolumn#1{m{#1}}
\def\imagetop#1{\vtop{\null\hbox{#1}}}
\usepackage{adjustbox}
\usepackage{subfig}
\usepackage{floatrow} % sidecapfloat
\usepackage{setspace}
\usepackage{xspace}
\usepackage{noReferences}
\usepackage{caption}
\usepackage{tikz}
\usepackage[customcolors]{hf-tikz}
%\usepackage{subcaption}
\usepackage{epsf}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{dsfont}
\usepackage{enumitem}
\usepackage{float}
\usepackage{cases}
\usepackage{verbatim}
\usepackage{array}
\usepackage{graphicx}
\usepackage{graphics}
\usepackage{multirow}
\usepackage{listings}
%\usepackage[authoryear]{natbib}
\usepackage{authblk}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{eqparbox}
\usepackage{xcolor}
\usepackage[backend=biber, citestyle=authoryear, maxcitenames=2, maxbibnames=2]{biblatex}
\usepackage{lipsum}
\usepackage{svg}

\bibliography{../biblio.bib}

\input{style}
\input{../mathdef}

% Removed from mathdef because of conflict with paper lcns class
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{remark}{Remark}
\newtheorem{property}{Property}

\begin{document}

%\background{ % Set the background to an image (background.pdf)
%\begin{tikzpicture}[remember picture,overlay]
%\draw (current page.north west)+(-2em,2em) node[anchor=north west]
%{\includegraphics[height=1.1\textheight]{background.png}};
%\end{tikzpicture}
%}
\begin{poster}{
grid=false,
borderColor=bordercol, % Border color of content boxes
headerColorOne=headercol1, % Background color for the header in the content boxes (left side)
headerColorTwo=headercol2, % Background color for the header in the content boxes (right side)
headerFontColor=headerfontcol, % Text color for the header text in the content boxes
boxColorOne=boxcolor, % Background color for the content in the content boxes
headershape=roundedright, % Specify the rounded corner in the content box headers
%headerfont=\Large\sf\bf, % Font modifiers for the text in the content box headers
headerfont=\Large\bf\textsc, %Sans Serif
textborder=rectangle,
background=none,
headerborder=open, % Change to closed for a line under the content box headers
boxshade=plain,
textfont={\setlength{\parindent}{0.0em}\sffamily},
headerheight={0.05\textheight},
eyecatcher=true
%columns=5
}
%
%----------------------------------------------------------------------------------------
%	Title and authors
%----------------------------------------------------------------------------------------
%
{
\includegraphics[width=14em]{./img/renault_group}
}
{
Practical Open-Loop Optimistic Planning
}
{
Edouard Leurent, Odalric-Ambrym Maillard
\vspace{-4\baselineskip}
}
{
\includegraphics[width=14em]{./img/inria_sc}
}

\setlength{\colheight}{0.92\textheight}

%----------------------------------------------------------------------------------------
%	Motivation
%----------------------------------------------------------------------------------------

\headerbox{\textsc{Motivation}}{name=motivation,span=1,column=0,row=0}{

We consider the optimal control of an MDP $\cM = (\cS, A, R, T, \gamma)$ with bounded rewards $R\in[0,1]$
\begin{itemize}[nolistsep]
    \item[\incarrow] $R$ and $T$ are \hlr{unknown}
    \item[\incarrow] Access to a \hlb{generative model} $s' \sim \probability{s'|s,a}$ and $r\sim\probability{r|s, a}$
    \item[\incarrow] \hlr{Fixed-budget} setting: the generative model is \hlr{costly}, can only be queried $n$ times
\end{itemize}
\texttt{UCT}: \hlr{doubly-exponential} $\xrightarrow{}$ \OPD: \hlg{polynomial}, \hlr{deterministic}
}

%----------------------------------------------------------------------------------------
%	Setting
%----------------------------------------------------------------------------------------
\headerbox{\textsc{Open-Loop Optimistic Planning}}{name=setting,span=1,column=0,row=0,below=motivation}{
\OLOP algorithm introduced in [\cite{Bubeck2010}].

%Split the budget $n$ into $M$ sequences of length $L$.
%Let $M$ be the largest integer such that $M \log M/(2 \log 1/\gamma) \leq n$\;
%Let $L = \log M / (2 \log 1/\gamma)$\;
\begin{enumerate}
    \item Sample $M$ sequences of actions of fixed length $L$
    \begin{center}
\includesvg[width=0.75\textwidth]{img/olop-explain.svg}
\end{center}
    \item Use the \hlb{return structure} to \hlb{generalise} to unseen sequences
    \begin{center}
\includesvg[width=0.75\textwidth]{img/olop-explain-2.svg}
\end{center}
\item Be \hlb{Optimistic in the Face of Uncertainty}\\
    \incarrow~ in \hlr{observed} and \hlr{future} rewards
\end{enumerate}

\begin{center}
\begin{algorithm}[H]
\DontPrintSemicolon
\footnotesize
\For{each episode $m = 1, \cdots, M$}{
Compute $U_a(m-1)$ from \eqref{eq:Ua} for all $a\in\mathcal{T}$\;
Compute $B_a(m-1)$ from \eqref{eq:Ba} for all $a\in A^L$\;\label{alg:b_values_compute}
Sample a sequence with highest B-value: $a^m \in \argmax_{a\in A^L} B_a(m-1)$.\;
}
\Return the most played sequence $a(n) \in \argmax_{a\in A^L} T_a(M)$
\caption{General structure for Open-Loop Optimistic Planning}
\label{algo:kl-olop}
\end{algorithm}
\end{center}

}

%----------------------------------------------------------------------------------------
% Tools
%----------------------------------------------------------------------------------------
\headerbox{\textsc{What's wrong with OLOP?}}
{name=tools,column=0,below=setting,span =1}
{

\hlr{Overly optimistic}, especially in the low-budget regime.

    \begin{equation}
         \underbrace{U^{\mu}_a(m)}_{\text{Upper bound}} \eqdef \underbrace{\hat{\mu}_a(m)}_{\text{Empirical mean}} + \underbrace{\sqrt{\frac{2 \log M}{T_a(m)}}}_{\text{Confidence interval}}
    \end{equation}
    \begin{equation}
    \label{eq:Ua}
        U_a(m) \eqdef \sum_{t=1}^h \underbrace{\gamma^t U^{\mu}_{a_{1:t}}(m)}_{\text{Past rewards}} + \underbrace{\frac{\gamma^{h+1}}{1-\gamma}}_{\text{Future rewards}}
    \end{equation}
    \begin{equation}
    \label{eq:Ba}
        B_a(m) \eqdef \inf_{1 \leq t \leq L} U_{a_{1:t}}(m)
    \end{equation}
    Intuitive explanation: 
    \begin{itemize}[noitemsep]
        \item Unintended behaviour happens when $\hler{U^{\mu}_a(m) > 1}, \forall a$. 
        \item Then the sequence $(U_{a_{1:t}}(m))_t$ is non-decreasing
        \item Then \hlr{$B_a(m) = U_{a_{1:1}}(m)$}
    \end{itemize}

\begin{center}
\includesvg[width=0.75\textwidth]{img/olop-explain-3.svg}    
\end{center}

\begin{center}
\texttt{OLOP} behaves as \hlr{uniform planning}!    
\end{center}
}

%----------------------------------------------------------------------------------------
%	KL-OLOP
%----------------------------------------------------------------------------------------
\headerbox{\textsc{Kullback-Leibler OLOP}}{name=klolop,column=1,span=2}{

\begin{minipage}[]{0.59\textwidth}
 We summon the upper-confidence bound from \texttt{kl-UCB} [\cite{Cappe2013}]:
    \begin{equation*}
        U^{\mu}_a(m) \eqdef \max \left\{q\in I: d(\hat{\mu}_a(m), q) \leq \frac{f(m)}{T_a(m)} \right\}
    \end{equation*}

    \begin{center}
    \begin{tabular}{ccc}
    \toprule
        Algorithm & \texttt{OLOP} & \texttt{KL-OLOP} \\
        \midrule
        Interval $I$ & $\mathbb{R}$ & [0, 1] \\
        Divergence $d$ & $d_{\texttt{QUAD}}$ & $d_{\texttt{BER}}$ \\
        $f(m)$ & $4 \log M$ & $2\log M + 2 \log\log M$\\
        \bottomrule
    \end{tabular}
    \end{center}
    with
    \begin{align*}
    d_{\texttt{QUAD}}(p,q) &\eqdef 2(p-q)^2\\
    d_{\texttt{BER}}(p, q) &\eqdef p \log \frac{p}{q} + (1-p)\log\frac{1-p}{1-q}
    \end{align*}
\end{minipage}
\begin{minipage}[]{0.4\textwidth}
    \includegraphics[width=\textwidth]{../img/ukl}
\end{minipage}

Conversely,
\begin{itemize}
    \item \hlg{$U^{\mu}_a(m) \in I = [0, 1], \forall a$}. 
    \item The sequence $(U_{a_{1:t}}(m))_t$ is \hlg{non-increasing}
    \item $B_a(m) = U_a(m)$, the \emph{bound sharpening} step is \hlb{superfluous}.
\end{itemize}
}

%----------------------------------------------------------------------------------------
%	Sample Complexity
%----------------------------------------------------------------------------------------
\headerbox{Sample complexity}{name=sample,span=2,column=1,below=klolop}{ 
\begin{theorem}[Sample complexity]
\label{thm:regret}
\KLOLOP enjoys the same asymptotic regret bounds as \OLOP. More precisely, \KLOLOP satisfies:
\begin{equation*}
    \expectedvalue r_n = \begin{cases}
      \tilde{0}\left(n^{-\frac{\log 1/\gamma}{\log \kappa'}}\right), & \text{if}\ \gamma\sqrt{\kappa'} > 1 \\
      \tilde{0}\left(n^{-\frac{1}{2}}\right), & \text{if}\ \gamma\sqrt{\kappa'} \leq 1
    \end{cases}
\end{equation*}
\end{theorem}
}

%----------------------------------------------------------------------------------------
%	Time and Memory Complexity
%----------------------------------------------------------------------------------------
\headerbox{Time and memory complexity}{name=time,span=2,column=1,below=sample}{

\begin{tabular}{>{\centering}p{0.45\textwidth}>{\centering}p{0.5\textwidth}}
\begin{minipage}[]{0.45\textwidth}

\textbf{Original \KLOLOP}

\emph{Compute $B_a(m-1)$ from \eqref{eq:Ba} \hlr{for all $a\in A^L$}}

\textbf{Lazy \KLOLOP}
\begin{center}
\includegraphics[width=0.8\textwidth]{../img/tree_svg-tex}    
\end{center}

\begin{theorem}[Consistency]
\label{thm:consistency}
Algorithm \ref{algo:lazy-kl-olop} is \hlg{identical} to Algorithm \ref{algo:kl-olop}.
\end{theorem}
\end{minipage}

&

\begin{minipage}{0.5\textwidth}
    \begin{algorithm}[H]
\DontPrintSemicolon
\footnotesize
Let $\Tau_0^+ = \LL_0^+ = \{\emptyset\}$\;
\For{each episode $m = 1, \cdots, M$}{
Compute $U_a(m-1)$ from \eqref{eq:Ua} for all \hlg{$a\in\Tau_{m-1}^+$}\;
Compute $B_a(m-1)$ from \eqref{eq:Ba} for all \hlg{$a\in \LL_{m-1}^+$}\;
Sample a sequence with highest B-value: $a \in \argmax_{\hleg{a\in \LL_{m-1}^+}} B_a(m-1)$\;
Choose an arbitrary continuation $a^m \in aA^{L-|a|}$\tcp*[h]{e.g. uniformly}
Let $\Tau_m^+ = \Tau_{m-1}^+$ and $\LL_m^+ = \LL_{m-1}^+$\;
\For{$t=1, \cdots, L$}{
    \If{$a^m_{1:t} \not \in \Tau_{m}^+$}{
    \hlb{Add $a^m_{1:t-1}A$ to $\Tau_{m}^+$ and $\LL_{m}^+$}\;
    \hlb{Remove $a^m_{1:t-1}$ from $\LL_{m}^+$}
    }
}
}
\Return the most played $a(n) \in \argmax_{a\in \LL_m^+} T_a(M)$
\caption{Lazy Open Loop Optimistic Planning}
\label{algo:lazy-kl-olop}
\end{algorithm}
\end{minipage}
\end{tabular}

\begin{property}[Time and memory complexity]
\begin{equation*}
    \frac{C(\texttt{Lazy KL-OLOP})}{C(\texttt{KL-OLOP})} = \frac{\hleg{nA}}{\hler{A^{L}}}
\end{equation*}
\end{property}

}

%
%%----------------------------------------------------------------------------------------
%% Experiments
%%----------------------------------------------------------------------------------------
%
\headerbox{Experiments}{name=experiments,span=2,column=1,below=time, above=bottom}{ 
     \includesvg[width=0.33\textwidth]{../img/hw_return}
     \includesvg[width=0.33\textwidth]{../img/gw_return}
     \includesvg[width=0.33\textwidth]{../img/gw_stoch_return}
     \begin{center}
         \textbf{Average return} over 100 runs –- along with its 95\% confidence interval –- with respect to the available budget $n$
     \end{center}

    \includesvg[width=0.33\textwidth]{../img/tree_OPD}    \includesvg[width=0.33\textwidth]{../img/tree_OLOP}
    \includesvg[width=0.33\textwidth]{../img/tree_KL-OLOP}
    \begin{center}
        \textbf{Expanded trees} for a budget $n=10^3$
    \end{center}
}

%----------------------------------------------------------------------------------------
%	Acknowledgements
%----------------------------------------------------------------------------------------
% \headerbox{Acknowledgements}{name=ack,column=0,span=1,below=tools}{
% This work has been supported by CPER Nord-Pas de Calais/FEDER DATA Advanced data science and technologies 2015-2020, the French Ministry of Higher Education and Research, INRIA, and the French Agence Nationale de la Recherche (ANR).
% }

%----------------------------------------------------------------------------------------
%	References
%----------------------------------------------------------------------------------------
\headerbox{References}{name=refs,column=0,span=1,below=tools, above=bottom}{
    {
    \AtNextBibliography{\footnotesize}
    \setlength{\bibitemsep}{3pt}
    \printbibliography[heading=none]
    }
}

\end{poster}

\end{document}